---
title: "Final Project 318"
author: "Marley & Kai"
date: "2022-11-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(usdm)
library(Metrics)
library(pROC)
library(caret)
library(bestglm)
library(rcompanion)
```

```{r upload dataset}
library("haven")
raw.data<-read_xpt("/cloud/project/P_RHQ.XPT") # does not include demographics data. 
```

# Central Variables & Codes:
SEQN - Respondent sequence number
RHD280 - Had a Hysterectomy? 
RHQ031 - Had regular periods in the past twelve months?
RHD043 - Reason not having regular periods?
RHQ076 - Seen a Doctor bc unable to become pregnant?
RHQ131 - Ever been pregnant?
RHQ160 - How many times have been pregnant?
RHD167 - Total number of deliveries
RHQ078 - Ever been treated for pelvic infection?
RHD180 - Age at first live birth?
RHD190 - Age at last live birth?
RHQ540 - Ever use female hormones?

```{r clarify data set to variables we are considering}
#colnames(raw.data)
data1<-raw.data[,c("SEQN", "RHD280", "RHQ031", "RHD043", "RHQ076", "RHQ131", "RHD143", "RHQ160", "RHD167", "RHQ078", "RHD180", "RHD190", "RHQ540")]
#head(data1, 4)
colnames(data1)<-c("ID", "hysterectomy", "regularPeriods", "periodReason", "doctor", "beenPregnant", "currentlyPregnant" , "numPregnant", "numDeliveries", "pelvicInfection", "ageFirstLiveBirth", "ageLastLiveBirth", "useFemaleHormones")
head(data1, 4)
```

```{r demographics data}
raw.demo<-read_xpt("/cloud/project/P_DEMO.XPT")
#head(raw.demo, 4)
```

# Central Demographic Variables to consider:
SEQN - Respondent sequence number
RIDRETH3 - Race/Hispanic origin w/ NH Asian
DMDBORN4 - Country of birth
DMDYRUSZ - Length of time in US
DMDEDUC2 - Education level - Adults 20+
DMDMARTZ - Marital status
SIALANG - Language of SP Interview
SIAINTRP - Interpreter used in SP Interview?
INDFMPIR - Ratio of family income to poverty
```{r relevant variables to demo}
demo.data1<-raw.demo[,c("SEQN","RIDAGEYR" ,"RIDRETH3", "DMDBORN4", "DMDYRUSZ", "DMDEDUC2", "DMDMARTZ", "SIALANG", "SIAINTRP", "INDFMPIR")]
#head(demo.data1, 4)
colnames(demo.data1)<-c("ID", "age", "race", "birthCountry", "timeInUS", "education", "maritalStatus", "interviewLanguage", "interpreter", "incomeToPovertyRatio")
head(demo.data1, 4)
```

```{r create trimmed dataset w hysterectomy IDs}
## first we should merge the demographics data and the hysterectomy data
## get the ID from the data1 dataset, find it in the demo data, and store the row value in a vector. 
data1$ID<-as.factor(data1$ID) #convert to factor
# create a new demographic dataset to proliferate with these selected values
demo.trimmed<-data.frame(matrix(ncol = ncol(demo.data1), nrow = 5314)) #5314 = total num observations in hysterectomy data
colnames(demo.trimmed)<-colnames(demo.data1)
IDs.handled<-0
for (i in 1:nrow(data1)) {
  IDs<-data1$ID # create list of IDs
  i.ID<- IDs[i]  #get the IDin position i
  #do all the ID's in the data1 dataset exist in the demo.data1 dataset
  pos.in.demo<- which(demo.data1$ID == i.ID) # find the index the ID exists in 
  #add the index to the new demo dataset
  demo.trimmed[i,]<-demo.data1[pos.in.demo,]
  IDs.handled=IDs.handled+1
}
```

```{r merge trimmed datasets}
# merge the two datasets - should have the same dimensions and therefore cbind will work. 
full.trimmed<- cbind(data1, demo.trimmed[,c(2:10)])

# then we trim the full dataset to only consider observations of hysterectomy with levels 'yes' and 'no'
yes<- which(full.trimmed$hysterectomy=='1')
no<- which(full.trimmed$hysterectomy=='2')
yes.trimmed<-full.trimmed[c(yes),]
no.trimmed<-full.trimmed[c(no),]
final.trimmed<-rbind(yes.trimmed, no.trimmed)
```

We want to explore the levels of each of the other variables, confirm they are of the correct type, and consider how missingness factors into our analysis. 
```{r data exploration and summary}
attach(final.trimmed)
summary(final.trimmed)
final.trimmed$ID<-as.factor(final.trimmed$ID)
final.trimmed$hysterectomy<-as.factor(final.trimmed$hysterectomy)
levels(as.factor(hysterectomy))
```



```{r deleting refused and don't know} 
#gets index of rows to be removed because refused or dont know
rows.remove <- which(periodReason == 99 | doctor == 9 | beenPregnant == 9 | numPregnant == 77 | numDeliveries == 77 | numDeliveries == 99 | pelvicInfection == 7 | pelvicInfection == 9 | ageFirstLiveBirth == 777 | ageFirstLiveBirth == 999 | ageLastLiveBirth == 999 | useFemaleHormones == 7 | useFemaleHormones == 9 | birthCountry == 77 | birthCountry == 99 | timeInUS == 77 | timeInUS == 99 | education == 7 | education == 9 | maritalStatus == 77 | maritalStatus == 99 | currentlyPregnant == 7 | currentlyPregnant == 9)

#creates new data set with rows removed
new.trimmed<-final.trimmed[-c(rows.remove),]
#deleted 99 rows now at 3931
```

```{r regular periods}
#12-150 years old
#no missing
new.trimmed[c(which(new.trimmed$regularPeriods==1)),3]<-"Yes"
new.trimmed[c(which(new.trimmed$regularPeriods==2)),3]<-"No"
```

```{r period reason}
#20-150 years old
(sum(is.na(new.trimmed$periodReason)/3917))*100 #44% missing
new.trimmed[is.na(new.trimmed$periodReason)==TRUE,]
#doesn't seem correlated with age, perhaps exclude 
```
```{r doctor}
#18-59 
(sum(is.na(new.trimmed$doctor)/3917))*100 #36% missing
new.trimmed[c(which(new.trimmed$age>=59)),5]<-"Does not apply"
new.trimmed[c(which(new.trimmed$doctor==1)),5]<-"Yes"
new.trimmed[c(which(new.trimmed$doctor==2)),5]<-"No"
```

```{r been pregnant}
#20-150 years old
(sum(is.na(new.trimmed$beenPregnant)/3917))*100 #0.03% missing
new.trimmed<-new.trimmed[-c(256),]
new.trimmed[c(which(new.trimmed$beenPregnant==1)),6]<-"Yes"
new.trimmed[c(which(new.trimmed$beenPregnant==2)),6]<-"No"
```

```{r currently pregnant}
#20-44
new.trimmed[c(which(new.trimmed$currentlyPregnant==1)),7]<-"Yes"
new.trimmed[c(which(new.trimmed$currentlyPregnant==2)),7]<-"No"
new.trimmed[c(which(new.trimmed$age>44)),7]<-"Does Not Apply"
new.trimmed[c(which(new.trimmed$age<20)),7]<-"Does Not Apply"
new.trimmed[is.na(new.trimmed$currentlyPregnant)==TRUE,]
```

```{r num pregnant}
#20-150 years old
(sum(is.na(new.trimmed$numPregnant)/3917))*100 #16% missing
new.trimmed[c(which(new.trimmed$beenPregnant=="No")),8]<-"Does not apply"
```

```{r num deliveries}
#12-150 years old
(sum(is.na(new.trimmed$numDeliveries)/3917))*100 #17% missing
new.trimmed[c(which(new.trimmed$beenPregnant=="No")),9]<-"Does not apply"
new.trimmed[c(which(new.trimmed$currentlyPregnant=="Yes")),9] <- "Does not apply"
new.trimmed[c(which(new.trimmed$numPregnant==1)),9] <- "Does not apply"
new.trimmed<-new.trimmed[-c(678,1118,2818),]
```

```{r pelvic infection}
#18-59 year olds
(sum(is.na(new.trimmed$pelvicInfection)/3917))*100 #35% missing
new.trimmed[c(which(new.trimmed$age>=59)),10]<-"Does not apply"
new.trimmed[c(which(new.trimmed$pelvicInfection==1)),10]<-"Yes"
new.trimmed[c(which(new.trimmed$pelvicInfection==2)),10]<-"No"
```

```{r age first live birth}
#20-59 year olds
(sum(is.na(new.trimmed$ageFirstLiveBirth)/3917))*100 #53% missing
new.trimmed[c(which(new.trimmed$age>=59 | new.trimmed$beenPregnant=="No" | new.trimmed$numDeliveries == 0)),11]<-"Does not apply"
new.trimmed[c(which(new.trimmed$currentlyPregnant=="Yes" & new.trimmed$numPregnant == 1)),11] <- "Does not apply"
missing.first.birth<-which(is.na(new.trimmed$ageFirstLiveBirth))
new.trimmed<-new.trimmed[-c(missing.first.birth),]
#31 people left, possibly impute agelastlivebirth if one delivery
```

```{r age last live birth}
#20-59 year olds
(sum(is.na(new.trimmed$ageLastLiveBirth)/3917))*100 #52% missing
new.trimmed[c(which(new.trimmed$age>=59)),12]<-"Does not apply"
new.trimmed[c(which(new.trimmed$beenPregnant=="No")),12]<-"Does not apply"
new.trimmed[c(which(new.trimmed$numDeliveries==0)),12]<-"Does not apply"
new.trimmed[c(which(new.trimmed$currentlyPregnant=="Yes" & new.trimmed$numPregnant == 1)),12] <- "Does not apply"
new.trimmed[is.na(new.trimmed$ageLastLiveBirth)==TRUE,]
missing.last.birth<-which(is.na(new.trimmed$ageLastLiveBirth))
new.trimmed<-new.trimmed[-c(missing.last.birth),]
```

```{r use female hormones}
#20-150 year olds
(sum(is.na(new.trimmed$useFemaleHormones)/3917))*100 #0.1% missing
new.trimmed<-new.trimmed[-c(678,1119,2829),]
new.trimmed[c(which(new.trimmed$useFemaleHormones==1)),13]<-"Yes"
new.trimmed[c(which(new.trimmed$useFemaleHormones==2)),13]<-"No"
levels(as.factor(new.trimmed$useFemaleHormones))
```

```{r age}
#0-150 year olds
```

```{r race}
#0-150 years old
new.trimmed[c(which(new.trimmed$race==1)),15]<-"Mexican American"
new.trimmed[c(which(new.trimmed$race==2)),15]<-"Other Hispanic"
new.trimmed[c(which(new.trimmed$race==3)),15]<-"Non-Hispanic White"
new.trimmed[c(which(new.trimmed$race==4)),15]<-"Non-Hispanic Black"
new.trimmed[c(which(new.trimmed$race==6)),15]<-"Non-Hispanic Asian"
new.trimmed[c(which(new.trimmed$race==7)),15]<-"Other Race-Including Multi-Racial"
```

```{r birth country}
#0-150 years old
new.trimmed[c(which(new.trimmed$birthCountry==1)),16]<-"US"
new.trimmed[c(which(new.trimmed$birthCountry==2)),16]<-"Other"
```

```{r time in us}
#0-150 years old
(sum(is.na(new.trimmed$timeInUS)/3917))*100 #73% missing
new.trimmed[c(which(new.trimmed$birthCountry=="US")),17]<-"Does Not Apply"
```

```{r education}
#20-150 years old
new.trimmed[c(which(new.trimmed$education==1)),18]<-"Less than 9th Grade"
new.trimmed[c(which(new.trimmed$education==2)),18]<-"9-11th Grade"
new.trimmed[c(which(new.trimmed$education==3)),18]<-"High School Graduate"
new.trimmed[c(which(new.trimmed$education==4)),18]<-"Some College or AA"
new.trimmed[c(which(new.trimmed$education==5)),18]<-"College Graduate or Above"
```

```{r marital status}
#20-150 years old
new.trimmed[c(which(new.trimmed$maritalStatus==1)),19]<-"Married/Living with Partner"
new.trimmed[c(which(new.trimmed$maritalStatus==2)),19]<-"Widowed, Divorced, or Separated"
new.trimmed[c(which(new.trimmed$maritalStatus==3)),19]<-"Never Married"
```

```{r interview language}
#0-150 years old
new.trimmed[c(which(new.trimmed$interviewLanguage==1)),20]<-"English"
new.trimmed[c(which(new.trimmed$interviewLanguage==2)),20]<-"Spanish"
```

```{r interpreter}
#0-150 years old
new.trimmed[c(which(new.trimmed$interpreter==1)),21]<-"Yes"
new.trimmed[c(which(new.trimmed$interpreter==2)),21]<-"No"
```

```{r income poverty ratio}
#0-150 years old
(sum(is.na(new.trimmed$incomeToPovertyRatio)/3917))*100 #0.1% missing
length(which(is.na(new.trimmed$incomeToPovertyRatio)==TRUE))
#look back at this still 472 missing
head(new.trimmed,6)
```
```{r hysterectomy labels}
#0-150 years old
new.trimmed$hysterectomy<- as.character(new.trimmed$hysterectomy)
new.trimmed[c(which(new.trimmed$hysterectomy=='1')),2]<- "Yes"
new.trimmed[c(which(new.trimmed$hysterectomy=='2')),2]<- "No"
#look back at this still 472 missing
head(new.trimmed)
```

```{r final data set}
data.final<-new.trimmed[,-c(1,3,4,7,22)] # remove ID, regular period, periodReason, currently pregnant, and incomeToPoveryRatio
summary(data.final)

attach(data.final)
```

```{r fix levels}
#table(data.final$hysterectomy)
#table(data.final$regularPeriods)
#table(data.final$doctor)
#table(data.final$beenPregnant)
#combine 8 - 11 as 8 to indicate 8+
data.final[c(which(data.final$numPregnant==8)),4]<-8
data.final[c(which(data.final$numPregnant==9)),4]<-8
data.final[c(which(data.final$numPregnant==10)),4]<-8
data.final[c(which(data.final$numPregnant==11)),4]<-8
table(data.final$numPregnant)
#table(data.final$numDeliveries)
#table(data.final$pelvicInfection)
table(data.final$ageFirstLiveBirth) #ages 40+ set as 40
data.final[c(which(data.final$ageFirstLiveBirth==40)),7]<-40
data.final[c(which(data.final$ageFirstLiveBirth==41)),7]<-40
data.final[c(which(data.final$ageFirstLiveBirth==42)),7]<-40
data.final[c(which(data.final$ageFirstLiveBirth==43)),7]<-40

table(data.final$ageLastLiveBirth) #ages 40+ set as 40
data.final[c(which(data.final$ageLastLiveBirth==40)),8]<-40
data.final[c(which(data.final$ageLastLiveBirth==41)),8]<-40
data.final[c(which(data.final$ageLastLiveBirth==42)),8]<-40
data.final[c(which(data.final$ageLastLiveBirth==43)),8]<-40
data.final[c(which(data.final$ageLastLiveBirth==44)),8]<-40
data.final[c(which(data.final$ageLastLiveBirth==45)),8]<-40

#table(data.final$useFemaleHormones)
#table(data.final$age)
#table(data.final$race)
#table(data.final$birthCountry)
#table(data.final$timeInUS)
#table(data.final$education)
#table(data.final$maritalStatus)
#table(data.final$interviewLanguage)
#table(data.final$interpreter)
```

```{r correct column type}
# age needs to be a factor
colnames(new.data)
col.types<-c()
for(i in 1:ncol(new.data)){
  type<-typeof(new.data[,i])
  col.types<-append(col.types,type)
}
col.types
# ensure that predictors are of type factor not character
data.final$hysterectomy<-as.factor(data.final$hysterectomy)
data.final$doctor<-as.factor(data.final$doctor)
data.final$beenPregnant<-as.factor(data.final$beenPregnant)
data.final$numPregnant<-as.factor(data.final$numPregnant)
data.final$numDeliveries<-as.factor(data.final$numDeliveries)
data.final$pelvicInfection<-as.factor(data.final$pelvicInfection)
data.final$ageFirstLiveBirth<-as.factor(data.final$ageFirstLiveBirth)
data.final$ageLastLiveBirth<-as.factor(data.final$ageLastLiveBirth)
data.final$useFemaleHormones<-as.factor(data.final$useFemaleHormones)
#data.final$age<- as.factor(data.final$age)
data.final$race<- as.factor(data.final$race)
data.final$birthCountry<-as.factor(data.final$birthCountry)
data.final$timeInUS<-as.factor(data.final$timeInUS)
data.final$education<-as.factor(data.final$education)
data.final$maritalStatus<-as.factor(data.final$maritalStatus)
data.final$interviewLanguage<-as.factor(data.final$interviewLanguage)
data.final$interpreter<-as.factor(data.final$interpreter)
```

```{r cramer v correlation}
new.data<-data.final[,-c(2, 3, 5, 12, 16, 17)]
#checking manually 
levels(data.final$hysterectomy)
  #no strongly correlated variables 
levels(data.final$doctor)
  #strongly correlated with pelvic infection - 0.70716319
# strongly correlated with age at first live birth - 0.54780
# strongly correlated w age at last live birth - 0.53040
# strongly corelated with age - 0.71650
levels(data.final$numPregnant) 
  #strongly correlated with been pregnant - 1.00000
  # strongly correlated with num deliveries - 0.56870 #remove numDeliveries. 
levels(data.final$beenPregnant) # remove been Pregnant
# strongly correlated with numpregnant - 1.0000
#strongly correlated with numdeliveries - 0.73210
# Strongly correlated with age -0.49000
levels(data.final$pelvicInfection)
#strongly correlated with doctor - 0.70716319
# strongly correlated with age at first live birth - 0.52850
#strongly correlated with age at last live birth - 0.52570
#strongly correlated with age - 0.71740
levels(data.final$useFemaleHormones)
  #no strongly correlated variables
levels(data.final$race)
  #strong correlation with interview language - 0.6191207
# strongly correlated with birth country - 0.70490
levels(data.final$birthCountry) # remove birthCountry
#linearly correlated with timeinUS - 1.00000
#strongly correlated with interview language - 0.50020
levels(data.final$timeInUS)
  #strongly correlated with interview language - 0.50730
levels(data.final$education)
  #somewhat atrongly correlated with interview language - 0.47180
levels(data.final$maritalStatus)
  #no strongly correlated variables
levels(data.final$interviewLanguage) # remove interview language 
  #no strongly correlated variables
levels(data.final$interpreter) # remove interpreter. 
  #no strongly correlated variables
  
attach(new.data)
correlation<-data.frame(matrix(ncol = ncol(new.data), nrow = ncol(new.data)))
colnames(correlation) <- colnames(new.data)
rownames(correlation)<- colnames(new.data)
for(i in 1:ncol(new.data)){
  for(j in 1:ncol(new.data)){
    cor<-cramerV(new.data[,i],new.data[,j])
    correlation[i,j]<-cor
  }
}
correlation
#cramerV(hysterectomy,doctor)
```

```{r all subset selection}
# X<-model.matrix(hysterectomy~.,data=new.data)[,-1]
# X<-X[,-34]
# Xy <- as.data.frame(X,hysterectomy)
# AIC.best<-bestglm(Xy,IC="AIC")
# AIC.best$BestModels
# BIC.best<-bestglm(new.data,IC="BIC",family=binomial)
# BIC.best$BestModel

#finding perfect linear dependency
# correlation<-data.frame(matrix(ncol = ncol(X), nrow = ncol(X)))
# colnames(correlation) <- colnames(X)
# rownames(correlation)<- colnames(X)
# for(i in 1:ncol(X)){
#   for(j in 1:ncol(X)){
#     cor<-cramerV(X[,i],X[,j])
#     correlation[i,j]<-cor
#   }
# }
# correlation
# 
# for(i in 1:75)
# {
#   for(j in (i+1):75)
#   {
#     if(correlation[i,j]==1)
#       print(c(i,j))
#   }
# }
# 
# correlation[34,58]
# X[,34]

# result.CP=leaps(x, y, int=TRUE, method=c("Cp"), nbest=10)
# which.min(result.CP$Cp)
# result.CP$which[71,]
# 
# result.adjr2=leaps(x, y, int=TRUE, method=c("adjr2"), nbest=10)
# which.min(result.adjr2$adjr2)
# result.adjr2$which[10,]
```
```{r all models}
model.full<-glm(hysterectomy~., data=new.data, family=binomial(link=logit))
result.BIC <- step(model.full,direction="both",k=log(dim(new.data)[1]),trace=0)
result.AIC <- step(model.full, direction="both",trace=0, k=2)
```

```{r k-fold}
#k fold cv
train_control <- trainControl(method = "cv", number = 10)

full <- train(hysterectomy~., data = new.data, trControl = train_control, method = "glm") # accuracy = 0.7989503
#full

model.AIC<-train(hysterectomy ~ numPregnant + ageLastLiveBirth + useFemaleHormones + age + race + education + maritalStatus, data=new.data, method="glm", trControl=train_control) # accuracy = 0.803953
#model.AIC

model.BIC<-train(hysterectomy ~ useFemaleHormones + age + education, data=new.data, method="glm", trControl=train_control) # accuracy = 0.7989454
#model.BIC

```
The k fold cross validation scores essentially tell us that each of the models performs comparably well alongside each other. The model with the greatest accuracy score (highest amount of correct classifications) is our stepwise regression AIC criterion model. This includes predictors:numPregnant, ageLastLiveBirth, useFemaleHormones, age, race, education, and maritalStatus. 


```{r final models}
full <- glm(hysterectomy~., data = new.data, family=binomial(link=logit)) 

aic<-glm(hysterectomy ~ numPregnant + ageLastLiveBirth + useFemaleHormones + age + race + education + maritalStatus, data=new.data, family=binomial(link=logit))

bic<-glm(hysterectomy ~ useFemaleHormones + age + education, data=new.data, family=binomial(link=logit))
```

```{r likelihood ratio}
library(lmtest)
lrtest(full,aic)
```

```{r ratio of hysterectomy}
table(new.data$hysterectomy)
845/(845+2946)
```

```{r full model performance metrics}
attach(new.data)
#threshold 0.1
pi.hat.full = predict(full,type="response") #predicted prob
Y.hat.full = ifelse(pi.hat>0.1,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)
plot(roc(hysterectomy,pi.hat.full))
auc(hysterectomy,pi.hat.full) # AUC: 0.8482
#threshold 0.15
pi.hat = predict(full,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.15,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)

#threshold 0.2
pi.hat = predict(full,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.2,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)

#threshold 0.25
pi.hat = predict(full,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.25,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)

#threshold 0.3
pi.hat = predict(full,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.3,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)

accuracy.1<-(1555+808)/(1555+808+37+1391)
specificity.1<-1555/(1555+1391)
sensitivity.1<-808/(808+37)
precision.1<-808/(808+1391)

accuracy.15<-(1827+763)/(1827+763+82+1119)
specificity.15<-1827/(1827+1119)
sensitivity.15<-763/(763+82)
precision.15<-763/(763+1119)

accuracy.2<-(2074+707)/(2074+707+138+872)
specificity.2<-2074/(2074+872)
sensitivity.2<-707/(707+138)
precision.2<-707/(707+872)

accuracy.25<-(2251+635)/(2251+635+210+695)
specificity.25<-2251/(2251+695)
sensitivity.25<-635/(635+210)
precision.25<-635/(635+695)

accuracy.3<-(2417+572)/(2417+572+273+529)
specificity.3<-2417/(2417+529)
sensitivity.3<-572/(572+273)
precision.3<-572/(572+529)

full.1<-c(accuracy.1, specificity.1, sensitivity.1, precision.1)
full.15<-c(accuracy.15, specificity.15, sensitivity.15, precision.15)
full.2<-c(accuracy.2, specificity.2, sensitivity.2, precision.2)
full.25<-c(accuracy.25, specificity.25, sensitivity.25, precision.25)
full.3<-c(accuracy.3, specificity.3, sensitivity.3, precision.3)
df.performance.full<-rbind(full.1, full.15, full.2, full.25, full.3)
colnames(df.performance.full)<-c('accuracy', 'specificity', 'sensitivity', 'precision')
rownames(df.performance.full)<-c('0.1', '0.15', '0.20', '0.25', '0.3')
df.performance.full
```



```{r AIC performance metrics}
#threshold 0.1
pi.hat.aic = predict(aic,type="response") #predicted prob
Y.hat.aic = ifelse(pi.hat>0.1,1,0) #dichomotize using 0.5
table(Y.hat.aic,hysterectomy)
plot(roc(hysterectomy,pi.hat.aic))
auc(hysterectomy,pi.hat.aic) # AUC: 0.846
#threshold 0.15
pi.hat = predict(aic,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.15,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)

#threshold 0.2
pi.hat = predict(aic,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.2,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)

#threshold 0.25
pi.hat = predict(aic,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.25,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)

#threshold 0.3
pi.hat = predict(aic,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.3,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)

accuracy.1<-(1544+808)/(1544+808+37+1402)
specificity.1<-1544/(1544+1402)
sensitivity.1<-808/(808+37)
precision.1<-808/(808+1402)

accuracy.15<-(1831+760)/(1831+760+85+1115)
specificity.15<-1831/(1831+1115)
sensitivity.15<-760/(760+85)
precision.15<-760/(760+1115)

accuracy.2<-(2083+705)/(2083+705+140+863)
specificity.2<-2083/(2083+863)
sensitivity.2<-705/(705+140)
precision.2<-705/(705+863)

accuracy.25<-(2265+629)/(2265+629+216+681)
specificity.25<-2265/(2265+681)
sensitivity.25<-629/(629+216)
precision.25<-629/(629+681)

accuracy.3<-(2405+560)/(2405+560+285+541)
specificity.3<-2405/(2405+541)
sensitivity.3<-560/(560+285)
precision.3<-560/(560+541)

aic.1<-c(accuracy.1, specificity.1, sensitivity.1, precision.1)
aic.15<-c(accuracy.15, specificity.15, sensitivity.15, precision.15)
aic.2<-c(accuracy.2, specificity.2, sensitivity.2, precision.2)
aic.25<-c(accuracy.25, specificity.25, sensitivity.25, precision.25)
aic.3<-c(accuracy.3, specificity.3, sensitivity.3, precision.3)
df.performance.aic<-rbind(aic.1, aic.15, aic.2, aic.25, aic.3)
colnames(df.performance.aic)<-c('accuracy', 'specificity', 'sensitivity', 'precision')
rownames(df.performance.aic)<-c('0.1', '0.15', '0.20', '0.25', '0.3')
df.performance.aic
```

```{r BIC performance metrics}
#threshold 0.1
pi.hat.bic = predict(bic,type="response") #predicted prob
Y.hat.bic = ifelse(pi.hat.bic>0.1,1,0) #dichomotize using 0.5
table(Y.hat.bic,hysterectomy)
plot(roc(hysterectomy,pi.hat.bic))
auc(hysterectomy,pi.hat.bic) # AUC: 0.8271 - all threshold have same area under the curve
#threshold 0.15
pi.hat = predict(bic,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.15,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)

#threshold 0.2
pi.hat = predict(bic,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.2,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)

#threshold 0.25
pi.hat = predict(bic,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.25,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)

#threshold 0.3
pi.hat = predict(bic,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.3,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)

accuracy.1<-(1462+797)/(1462+797+48+1484)
specificity.1<-1462/(1462+1484)
sensitivity.1<-797/(797+48)
precision.1<-797/(797+1484)

accuracy.15<-(1830+737)/(1830+737+108+1116)
specificity.15<-1830/(1830+1116)
sensitivity.15<-737/(737+108)
precision.15<-737/(737+1116)

accuracy.2<-(2112+659)/(2112+659+186+834)
specificity.2<-2112/(2112+834)
sensitivity.2<-659/(659+186)
precision.2<-659/(659+834)

accuracy.25<-(2291+585)/(2291+585+260+655)
specificity.25<-2291/(2291+655)
sensitivity.25<-585/(585+260)
precision.25<-585/(585+655)

accuracy.3<-(2445+505)/(2445+505+340+501)
specificity.3<-2445/(2445+501)
sensitivity.3<-505/(505+340)
precision.3<-505/(505+501)

bic.1<-c(accuracy.1, specificity.1, sensitivity.1, precision.1)
bic.15<-c(accuracy.15, specificity.15, sensitivity.15, precision.15)
bic.2<-c(accuracy.2, specificity.2, sensitivity.2, precision.2)
bic.25<-c(accuracy.25, specificity.25, sensitivity.25, precision.25)
bic.3<-c(accuracy.3, specificity.3, sensitivity.3, precision.3)
df.performance.bic<-rbind(bic.1, bic.15, bic.2, bic.25, bic.3)
colnames(df.performance.bic)<-c('Accuracy', 'Specificity', 'Sensitivity', 'Precision')
rownames(df.performance.bic)<-c('0.1', '0.15', '0.20', '0.25', '0.3')
df.performance.bic
```
```{r ROC curves}
par(mfrow=c(3,1))
plot(roc(hysterectomy,pi.hat.full))
plot(roc(hysterectomy,pi.hat.aic))
plot(roc(hysterectomy,pi.hat.bic))

df.performance<-rbind(full.2, aic.2, bic.2)
colnames(df.performance)<-c('Accuracy', 'Specificity', 'Sensitivity', 'Precision')
rownames(df.performance)<-c('Saturated Model', 'Stepwise AIC', 'Stepwise BIC')
df.performance
```

```{r model diagnostics}
plot(aic) 
#outliers at 4317, 1313, 1730, 1903, 3228, 4780
full.trimmed[c(4317, 1313, 1730, 1903, 3228, 4780),] #why are the first and last included if there's not even 4000 rows? 
summary(aic)
new.trimmed[c(which(new.trimmed$ID=="121891")),]
new.trimmed[c(which(new.trimmed$ID=="113109")),]
new.trimmed[c(which(new.trimmed$ID=="114320")),]
new.trimmed[c(which(new.trimmed$ID=="114803")),]
new.trimmed[c(which(new.trimmed$ID=="118679")),]
new.trimmed[c(which(new.trimmed$ID=="123273")),]

# rows<- c(which(new.trimmed$ID=="121891"),which(new.trimmed$ID=="113109"),which(new.trimmed$ID=="114320"),
#          which(new.trimmed$ID=="114803"),which(new.trimmed$ID=="118679"),which(new.trimmed$ID=="123273"))

devdf<-data.frame(matrix(ncol=2,nrow=nrow(new.data)))
devdf[,1]<-1:nrow(new.data)
for (i in 1:nrow(new.data)) 
{
  new.data2<-new.data[-c(i),]
  aic2<-glm(hysterectomy ~ numPregnant + ageLastLiveBirth + useFemaleHormones + age + race + education + maritalStatus, data=new.data2, family=binomial(link=logit))
  dev<-aic$deviance - aic2$deviance
  devdf[i,2]<-dev
}

plot(x=devdf[,1],y=devdf[,2],xlab="Case Number",ylab="Delta Deviance",main="Delta Deviance Plot")

devdf[c(which(devdf[,2]>=38)),]

new.data<-new.data[-c(1,5,52,254,272,510,770),]

aic<-glm(hysterectomy ~ numPregnant + ageLastLiveBirth + useFemaleHormones + age + race + education + maritalStatus, data=new.data, family=binomial(link=logit))
summary(aic)


#library(blorr)
#blr_plot_diag_difdev(aic)
```

Given the relatively small number of predators in the best (BIC) model, it's reasonable to explore possible two-way interaction terms.
```{r consider interaction terms to BIC model BIC criterion}
best.BIC <- glm(hysterectomy ~ useFemaleHormones + age + education, data=new.data, family=binomial(link=logit)) #define your best BIC glm model here
step(best.BIC, .~.^2, direction="both", k=log(dim(new.data)[1])) 
```

```{r consider interaction terms to BIC model AIC criterion}
step(best.BIC, .~.^2, direction="both", k=2) 
```

Best Interaction Model based on Best BIC first order model using stepwise regression with BIC as criterion:
hysterectomy ~ useFemaleHormones + age + education + useFemaleHormones:age

Best Interaction Model based on Best BIC first order model using stepwise regression with AIC as criterion:
hysterectomy ~ useFemaleHormones + age + education + useFemaleHormones:age

Using stepwise regression with AIC and BIC as criterion for selecting the best second order model based on our best BIC first order model produced the same result. hat(log(hysterectomy)) = B0 + B1*useFemaleHormones + B2*age + B3*education + B4*useFemaleHormones:age

We can conduct k fold cross validation for comparison
```{r cv on interaction}
model.BIC.Interact<-train(hysterectomy ~ useFemaleHormones + age + education + useFemaleHormones:age, data=new.data, method="glm", trControl=train_control) # accuracy = 0.8050647
```
In terms of accuracy our interaction model performs best but again only marginally.

We can then perform another likelihood ratio test to see if the addition of the interaction terms sufficiently fits the data better than the reduced model (in this case our first order AIC model). 
```{r likelihood ratio test interaction vs. AIC}
interaction<-glm(hysterectomy ~ useFemaleHormones + age + education + useFemaleHormones:age, data=new.data, family=binomial(link=logit))
library(lmtest)
lrtest(bic,interaction)
lrtest(aic, interaction)
```

```{r performance metrics}
#threshold 0.1
pi.hat = predict(interaction,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.1,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)
plot(roc(hysterectomy,pi.hat))
auc(hysterectomy,pi.hat) # AUC: 0.8266
#threshold 0.15
pi.hat = predict(interaction,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.15,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)

#threshold 0.2
pi.hat = predict(interaction,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.2,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)

#threshold 0.25
pi.hat = predict(interaction,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.25,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)

#threshold 0.3
pi.hat = predict(interaction,type="response") #predicted prob
Y.hat = ifelse(pi.hat>0.3,1,0) #dichomotize using 0.5
table(Y.hat,hysterectomy)

accuracy.1<-(1534+793)/(1534+793+52+1412)
specificity.1<-1534/(1534+1412)
sensitivity.1<-793/(793+52)
precision.1<-793/(793+1412)

accuracy.15<-(1834+731)/(1834+731+114+1112)
specificity.15<-1834/(1834+1112)
sensitivity.15<-731/(731+114)
precision.15<-731/(731+1112)

accuracy.2<-(2067+670)/(2067+670+175+879)
specificity.2<-2067/(2067+879)
sensitivity.2<-670/(670+175)
precision.2<-670/(670+879)

accuracy.25<-(2234+608)/(2234+608+237+712)
specificity.25<-2234/(2234+712)
sensitivity.25<-608/(608+237)
precision.25<-608/(608+712)

accuracy.3<-(2374+532)/(2374+532+313+572)
specificity.3<-2374/(2374+572)
sensitivity.3<-532/(532+313)
precision.3<-532/(532+572)

interact.1<-c(accuracy.1, specificity.1, sensitivity.1, precision.1)
interact.15<-c(accuracy.15, specificity.15, sensitivity.15, precision.15)
interact.2<-c(accuracy.2, specificity.2, sensitivity.2, precision.2)
interact.25<-c(accuracy.25, specificity.25, sensitivity.25, precision.25)
interact.3<-c(accuracy.3, specificity.3, sensitivity.3, precision.3)
df.performance.interact<-rbind(interact.1, interact.15, interact.2, interact.25, interact.3)
colnames(df.performance.interact)<-c('accuracy', 'specificity', 'sensitivity', 'precision')
rownames(df.performance.interact)<-c('0.1', '0.15', '0.20', '0.25', '0.3')
df.performance.interact
```
Our first order AIC model and first order full model still perform best in terms of sensitivity. 